---
title: "Lab-01-Deus."
author: "Stephanie Deus"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## STEP 1 - CREATE A NEW GAME
## This function sets up a new game.

```{r}

create_game <- function() {
    a.game <- sample(x = c("goat", "goat", "car"), size = 3, replace = FALSE)
    return(a.game)
}
```

```{r}
set.seed(123)
game_setup_1 <- create_game()
game_setup_2 <- create_game()
game_setup_3 <- create_game()
print(game_setup_1)
print(game_setup_2)
print(game_setup_3)

```
## STEP 2 - CONTESTANT SELECTS A DOOR
## The contestant makes their first selection. 
```{r}

select_door <- function() {
    doors <- c(1, 2, 3)
    a.pick <- sample(doors, size = 1)
    return(a.pick)
}

```

```{r}
initial_pick_1 <- select_door()
initial_pick_2 <- select_door()
initial_pick_3 <- select_door()
print(initial_pick_1)
print(initial_pick_2)
print(initial_pick_3)

```

## STEP 3 - HOST OPENS GOAT DOOR
## The host will always open a door with a goat behind it. But it can’t be a door the contestant has already selected. So it must be a door that is not a car and not a current contestant selection.


```{r}

open_goat_door <- function(game, a.pick) {
    goat_doors <- which(game == "goat" & seq_along(game) != a.pick)
    opened.door <- sample(goat_doors, size = 1)
    return(opened.door)
}
```


```{r}
this_game <- create_game()
initial_pick <- select_door()
opened_door <- open_goat_door(this_game, initial_pick)
print(this_game)
print(initial_pick)
print(opened_door)

```


## Function 4: change_door
## This function represents the contestant's game-playing strategy (stay or switch). If stay=TRUE, the contestant keeps their initial pick. If stay=FALSE, the contestant switches doors.
```{r}

change_door <- function(stay, opened.door, a.pick) {
    if (stay) {
        final.pick <- a.pick
    } else {
        remaining_door <- setdiff(1:3, c(a.pick, opened.door))
        final.pick <- sample(remaining_door, size = 1)
    }
    return(final.pick)
}

```

```{r}

opened_door_test <- open_goat_door(this_game, initial_pick)
final_pick_stay_test <- change_door(stay = TRUE, opened.door = opened_door_test, a.pick = initial_pick)
final_pick_switch_test <- change_door(stay = FALSE, opened.door = opened_door_test, a.pick = initial_pick)
print(opened_door_test)
print(final_pick_stay_test)
print(final_pick_switch_test)

```

## Function 5: determine_winner
## This function determines if the contestant has won or lost based on the final pick and the game setup.
```{r}
determine_winner <- function(final.pick, game) {
    print(final.pick)
    print(game)
  
    if (game[final.pick] == "car") {
        return("WIN")
    } else {
        return("LOSE")
    }
}
```

```{r}
final_pick_win_test <- change_door(stay = TRUE, opened.door = opened_door_test, a.pick = initial_pick)
final_pick_lose_test <- change_door(stay = FALSE, opened.door = opened_door_test, a.pick = initial_pick)
result_win_test <- determine_winner(final_pick_win_test, this_game)
result_lose_test <- determine_winner(final_pick_lose_test, this_game)
print(final_pick_win_test)
print(final_pick_lose_test)
print(result_win_test)
print(result_lose_test)
```





