---
title: "Lab 2"
author: "Stephanie Deus"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
play_game <- function()

#Step 1
create_game <- function()
{
    a.game <- sample( x=c("goat","goat","car"), size=3, replace=F )
    return( a.game )
} 
```

```{r}

#Step 2
select_door <- function()
{
  doors <- c(1,2,3) 
  a.pick <- sample( doors, size=1 )
  return( a.pick )  # number between 1 and 3
}
```

```{r}
#Step 3
open_goat_door <- function( game, a.pick )
{
   doors <- c(1,2,3)
   # if contestant selected car,
   # randomly select one of two goats 
   if( game[ a.pick ] == "car" )
   { 
     goat.doors <- doors[ game != "car" ] 
     opened.door <- sample( goat.doors, size=1 )
   }
   if( game[ a.pick ] == "goat" )
   { 
     opened.door <- doors[ game != "car" & doors != a.pick ] 
   }
   return( opened.door ) # number between 1 and 3
}
```

```{r}

#Step 4
change_door <- function( stay=T, opened.door, a.pick )
{
   doors <- c(1,2,3) 
   
   if( stay )
   {
     final.pick <- a.pick
   }
   if( ! stay )
   {
     final.pick <- doors[ doors != opened.door & doors != a.pick ] 
   }
  
   return( final.pick )  # number between 1 and 3
}
```

```{r}
#Step 5
determine_winner <- function( final.pick, game )
{
   if( game[ final.pick ] == "car" )
   {
      return( "WIN" )
   }
   if( game[ final.pick ] == "goat" )
   {
      return( "LOSE" )
   }
}
```

```{r}
play_game <- function( )
  
{
  new.game <- create_game()
  first.pick <- select_door()
  opened.door <- open_goat_door( new.game, first.pick )
  final.pick.stay <- change_door( stay=T, opened.door, first.pick )
  final.pick.switch <- change_door( stay=F, opened.door, first.pick )
  outcome.stay <- determine_winner( final.pick.stay, new.game  )
  outcome.switch <- determine_winner( final.pick.switch, new.game )
  

  # game.results <- bundle the results
  # return( <<< game.results >>> )
  
  strategy <- c("stay","switch")
  outcome <- c(outcome.stay,outcome.switch)
  game.results <- data.frame( strategy, outcome,
                              stringsAsFactors=F )
  return( game.results )
}

determine_winner <- function(final.pick, game) {
  ifelse(any(game[final.pick] == "car", na.rm = TRUE), "WIN", "LOSE")
}

play_game()
```



```{r}

results.df <- NULL   # collector

for( i in 1:10000 )  # iterator
{
  game.outcome <- play_game()
  # binding step
  results.df <- rbind( results.df, game.outcome )
}

table( results.df ) 


```

```{r}

game.outcome <- play_game()
results.df <- rbind( results.df, game.outcome )


```

```{r}

# creating an empty collector object
results.df <- NULL   
# later, inside the loop
results.df <- rbind( results.df, game.outcome )

```

```{r}
x <- 1:4
c( x, 9 )
```

```{r}
game1 <- play_game()
game2 <- play_game()
rbind( game1, game2 )
```
```{r}
c( game1, game2 )

```
```{r}
results.df <- NULL 
rbind( results.df, game.outcome )
```
```{r}
# Modified create_game function for 3 doors with 2 cars and 1 goat
create_game <- function() {
    a.game <- sample(x = c("goat", "car", "car"), size = 3, replace = FALSE)
    return(a.game)
}
```

```{r}

# Modified open_goat_door function to handle 3 doors
open_goat_door <- function(game, a.pick) {
    non_chosen_doors <- setdiff(1:3, a.pick)
    goat_door <- non_chosen_doors[game[non_chosen_doors] == "goat"]
    return(goat_door)
}
```

```{r}
# Modified play_game function for the 3-door setup
play_game <- function() {
    new.game <- create_game()
    first.pick <- select_door()
    goat_door <- open_goat_door(new.game, first.pick)
    
    # Assume contestant always switches for simplicity
    remaining_door <- setdiff(1:3, c(first.pick, goat_door))
    final.pick <- remaining_door
    outcome <- determine_winner(final.pick, new.game)
    
    strategy <- "switch"  # Contestant always switches
    game.results <- data.frame(strategy, outcome, stringsAsFactors = FALSE)
    return(game.results)
}

```

```{r}
# Simulation loop
results.df <- NULL


for (i in 1:10000) {
    game.outcome <- play_game()
    results.df <- rbind(results.df, game.outcome)
}

# Display results
table(results.df) %>% 
  prop.table(margin = 1) %>% 
  round(2)

```
**Q1: Is SWITCH still the dominant strategy?**

Yes, according to my data above switch is the the highest win.


Part II


```{r}

open_goat_door <- function(game_setup, contestant_pick) {
  goat_doors <- which(game_setup == "goat" & seq_along(game_setup) != contestant_pick)
  opened_door <- sample(goat_doors, 1)
  return(opened_door)
}

create_custom_game <- function(cars, goats, total_doors) {
  game_setup <- c(rep("goat", goats), rep("car", cars))
  shuffled_setup <- sample(game_setup, size = total_doors, replace = TRUE)  
  return(shuffled_setup)
}

play_game_custom <- function(cars, goats, total_doors) {
  new.game <- create_custom_game(cars, goats, total_doors)
  first.pick <- sample(1:total_doors, 1)  
  opened.door <- open_goat_door(new.game, first.pick)
  final.pick.stay <- change_door(stay = TRUE, opened.door, first.pick)
  final.pick.switch <- change_door(stay = FALSE, opened.door, first.pick)
  outcome.stay <- determine_winner(final.pick.stay, new.game)
  outcome.switch <- determine_winner(final.pick.switch, new.game)
  
  strategy <- c("stay", "switch")
  outcome <- c(outcome.stay, outcome.switch)
  game.results <- data.frame(strategy, outcome, stringsAsFactors = FALSE)
  return(game.results)
}
```

```{r}

# Simulation for 1 car, 9 goats
total_doors_scenario1 <- 10
results_scenario1 <- NULL

for (i in 1:10000) {
  game.outcome <- play_game_custom(cars = 1, goats = 9, total_doors = total_doors_scenario1)
  results_scenario1 <- rbind(results_scenario1, game.outcome)
}

# Simulation for 2 cars, 8 goats
total_doors_scenario2 <- 10
results_scenario2 <- NULL

for (i in 1:10000) {
  game.outcome <- play_game_custom(cars = 2, goats = 8, total_doors = total_doors_scenario2)
  results_scenario2 <- rbind(results_scenario2, game.outcome)
}

# Simulation for 3 cars, 7 goats
total_doors_scenario3 <- 10
results_scenario3 <- NULL

for (i in 1:10000) {
  game.outcome <- play_game_custom(cars = 3, goats = 7, total_doors = total_doors_scenario3)
  results_scenario3 <- rbind(results_scenario3, game.outcome)
}
```

```{r}
table_scenario1 <- table(results_scenario1$strategy, results_scenario1$outcome)
table_scenario2 <- table(results_scenario2$strategy, results_scenario2$outcome)
table_scenario3 <- table(results_scenario3$strategy, results_scenario3$outcome)

prop_table_scenario1 <- prop.table(table_scenario1, margin = 1)
prop_table_scenario2 <- prop.table(table_scenario2, margin = 1)
prop_table_scenario3 <- prop.table(table_scenario3, margin = 1)

round(prop_table_scenario1, 2)
round(prop_table_scenario2, 2)
round(prop_table_scenario3, 2)


```

**Q1: Is SWITCH still the dominant strategy?**

Scenario with 1 car and 9 goats: Stay: 10% Wins, Switch: 11% Wins
Scenario with 2 cars and 8 goats: Stay: 20% Wins, Switch: 22% Wins
Scenario with 3 cars and 7 goats: Stay: 31% Wins, Switch: 33% Wins

In all scenarios, the strategy of switching doors gives a higher proportion of wins compared to the strategy of staying with the initial choice. The chances of winning increase when participants choose to switch doors.


**Q2: How much better off are you switching in the original game? How much did it improve your chances of winning?**

In the original Monty Hall Problem with 3 doors, 1 car, and 2 goats, participants who choose to switch doors have a slightly better chance of winning (11%) compared to those who stay with their initial choice (10%). Switching improves the chances of winning by 1%.


**Q3: How much better off are you switching in each of the three scenarios with 10 doors and 1 to 3 cars?**

Scenario with 1 car and 9 goats: 1% better off switching
Scenario with 2 cars and 8 goats: 2% better off switching
Scenario with 3 cars and 7 goats: 2% better off switching

In each scenario, switching improves the chances of winning, and the improvement varies between 1% and 2%.




